---
title: "Autos Wrangling"
author: "COMP/STAT 112"
date: "2024-03-28"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r include = FALSE}
# This chunk just sets up some styling (eg: default size of our images)
knitr::opts_chunk$set(
  collapse = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.height = 3, 
  fig.width = 5,
  fig.align = 'center')
```

\
\

```{r}

library(tidyverse)
library(ISLR2)
```

\
\

# Wrangle a Dataset 

\

Our goal in this markdown is to wrangle the `Auto` dataset to make it more user friendly. 

Here is the default dataset as it comes via the `ISLR2` package.  

\
\


```{r}

head(Auto)


```

\
\

After looking at this table we decide we want to

1) break up the `name` variable into `make` and `model` and clean up any issues with that conversion, we will also move these new variables to the first two columns.

2) capitalize the `make` variable and the first letter of the `model` variable

3) make the `origin` variable into an appropriate character variable instead of 1-3.

4) shorten some of the variable names so that the displayed tables are more compact.

Here is our first pass at creating the `make` and `model` variables.

\
\


```{r}

str(Auto)

Auto1 <- Auto %>%
  mutate(temp=str_extract(name,"[a-z]+ ")) %>% # strip off characters up to the first space
  mutate(make=str_sub(name,1,nchar(temp)-1)) %>%
  mutate(model=str_sub(name,nchar(temp)+1,nchar(as.character(name)))) %>%
  select(make,model,mpg,name)

head(Auto1)

table(Auto1$make)

summary(nchar(Auto1$model))

Auto1 %>% filter(is.na(model))

```

\
\

## Create `Auto_temp`

\

From the output of the `table(Auto1$make)` command we notice that there are many typos, misspellings, or multiple names used for the `make` variable.  We also notice that there are some NAs for the `model` variable.  For now we will drop those rows with na.omit().


In the next code chunk create a temporary dataframe called `Auto_temp`.  Use the `mutate()` command and string commands to fix the problems noted above with the `make` variable.  Also, in the same code chunk recode the `origin` variable with the appropriate character strings.

\
\

```{r}

# create Auto_temp in this code chunk

Auto_temp <- Auto 


```


## Try again to create `make` and `model`

\

In the next code chunk we will try again to split the `name` variable (now for `Auto_temp`) into `make` and `model`.  We will also capitalize the first letter of the `make` variable and the first letter of the first word of the `model` variable.

\  
  
```{r}


# create Autos from Auto_temp in this code chunk

Autos <- Auto_temp
 

```


\
\

##  Simple Exercises

\
\

Use the more user-friendly `Autos` to make some tables and plots

\


### Exercise 1
\

Make a table of all European cars for the year 1980.

\
\


### Exercise 2

\

Make a table of the average miles per gallon by make for all makes with at least 5 cars in the table.  sort the table from largest to smallest in terms of average mpg.

\
\

### Exercise 3

\

Make a scatterplot of year (as a factor) versus mpg colored by the origin variable.


\





