---
title: "Day4_Practice"
author: "Brady Brunswig"
date: "2024-01-30"
output: html_document
---
```{r,message=FALSE}
library(tidyverse)
library(fivethirtyeight)
```
```{r}
str(mtcars)
names(mtcars)
```
```{r}

ggplot(mtcars,aes(x=qsec))+
  geom_histogram(color="white", fill="orange",bins=8)
```
exersise 2
```{r}

ggplot(mtcars,aes(x=as.factor(vs)))+
  geom_bar( fill="purple")
```
```{r}

ggplot(mtcars,aes(x=vs))+
  geom_bar( fill="purple")
```
it changes the x-axis to act as a quantitative variable and when comparing just 2 different variables it doesn't make much sense to have the axis show 0.5 and 1.5, etc

exersise 3
```{r}
str(elo_blatter)
names(elo_blatter)

```
209 rows, 8 columns
Country is catergorical? elo98,elo15,popu06,gdp06 are all quanatative, and confederation,popu_source,and gdp_source are all catergorical


```{r}

ggplot(elo_blatter,aes(x=elo98))+
  geom_histogram(color="white", fill="black",bins=20)
```
Elo98 is more or less normally distibuted along the x axis, with an even amount of outliers on each side

```{r}

ggplot(elo_blatter,aes(x=confederation))+
  geom_bar(color="white", fill="black")
```

UEFA seems to be the most frequent confederation, along with afc and caf. 