---
title: "Day4 Bivariate Viz Plots"
author: "STAT/COMP 112, secs. 2 and 3"
date: "2024-01-30"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Code for the Bivariate Quantitative Plots


Below is the code which produced the scatterplots displayed in the  Class notes for Day 4.

\


First, an unadorned scatterplot of the elo data

\

```{r,message=FALSE}
library(tidyverse)
library(fivethirtyeight)

ggplot(elo_blatter,aes(x=elo98,y=elo15)) +
  geom_point()
```

\

Here is the same plot with the point representing USA highlighted

Note: we have also labelled the point.  This was done using the package **gghighlight**.

\

```{r}
library(gghighlight)

ggplot(elo_blatter,aes(x=elo98,y=elo15)) +
  geom_point(color="red",size=2) +  
  gghighlight(country=="United States",label_key = country,
              unhighlighted_params = list(color="blue",size=1))

```


# Code for the Bivariate Categorical Plots

\


Below is the code which produced the barplots displayed in the notes for Day 4.

\

```{r}
library(ISLR2)

str(Auto)
names(Auto)

## Note in the commands below, in the aesthetic we use the option 
## fill = <second variable>     to make the colored bar segments

ggplot(Auto,aes(x=as.factor(origin),fill=as.factor(cylinders)))+
  geom_bar()

ggplot(Auto,aes(x=as.factor(cylinders),fill=as.factor(origin)))+
  geom_bar()
```

\


# Day 4 Exercises

\

Create a new .rmd file called **Day 4 Exercises** and do these problems in that markdown document

\

in the first code chunk of your document you will need to load in the packages containing the data. so run

**library(fivethirtyeight)**

and

**library(ISLR2)**

\

## Exercise 1

\

Using code similar to the highlighted scatterplot for the elo rankings, replace the option 

*country = "United States"*  with  *confederation = "CONMEBOL"*

also replace  *size = 2* with *size = 3* in the geom_point() options

Explain what these changes do to the plot.

Now do the same thing for the confederation *CONCACAF* (this is the federation to which the USA belongs).

Describe similarities and differences that you see between the elo rankings of the countries in the two federations.

Also, how is the second plot different from the first, and can you explain why it is different?
\
```{r}
ggplot(elo_blatter,aes(x=elo98,y=elo15)) +
  geom_point(color="red",size=3) +  
  gghighlight(confederation=="CONCACAF",label_key = country,
              unhighlighted_params = list(color="blue",size=1))

```
\
this outlines all the countries in the confederation of conmebol. also changed the size of the dots

Too many points are highlighted to have the text on the graph
## Exercise 2

\

In the ISLR2 repository of data there is a dataset called **BrainCancer** which has 5 categorical variables with each with four or fewer categories.  Using different pairs of variables make *at least three* segmented barplots using these variables and describe what these plots tell you about the relationships between the variables.

```{r}
ggplot(BrainCancer,aes(x=as.factor(stereo),fill=as.factor(diagnosis)))+
  geom_bar()
ggplot(BrainCancer,aes(x=as.factor(loc),fill=as.factor(diagnosis)))+
  geom_bar()
ggplot(BrainCancer,aes(x=as.factor(diagnosis),fill=as.factor(sex)))+
  geom_bar()
```

SRS seems to have a lower total diagnosis number than SRT screening

most of the cancers have a higher rate in supratertorial location vs infratentorial except for the other catergory

Females seem to have higher rates of cancer in Meninioma, but in all other categories the males seem to have higher rates when compared to women