---
title: "More Mapping (Spatial Visualization)"
author: "Dan O'Loughlin"
date: "2024-02-07"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r include = FALSE}
# This chunk just sets up some styling (eg: default size of our images)
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.height = 4,
  fig.width = 6,
  fig.align = 'center')
```

```{r}
## Load in pacakages

library(tidyverse)
library(rnaturalearth)

# Get info about country boundaries across the world
# in a "sf" or simple feature format
world_boundaries <- ne_countries(returnclass = "sf")


# Import starbucks location data

starbucks <- read.csv("https://bcheggeseth.github.io/112_fall_2023/data/starbucks.csv")


## Simple wrangle of Starbucks dataset to pull out MN Starbucks and drop some columns

MN_Starbucks<-starbucks%>%
  filter(State.Province=="MN")%>%
  select(Store.Number,Store.Name,Ownership.Type,Street.Address,City,Longitude,Latitude)
```

\

# Map of all the Minnesota Starbucks

NOTE:  The labeling here is not very effective!

\


```{r}
# Load packages
library(sf)
library(maps)

states <- map_data("state")

mn_data<-subset(states, region == "minnesota")

ggplot(data = mn_data, mapping = aes(x = long, y = lat)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")+ 
  geom_point(
    data = MN_Starbucks,
    aes(x = Longitude, y = Latitude, color=Ownership.Type),
    alpha = 1, size = 2)+
  geom_text(data=MN_Starbucks, aes(Longitude, Latitude, label=City))
  

```

\

# Map of some Minnesota Cities and Towns

NOTE:  Here labeling the points by the city is useful.

\


```{r}

City<-c("Rochester","Duluth","Minneapolis","Cambridge","Ely","Morris","Tofte")
Lat<-c(44.0,46.8,45.0,45.5,47.9,45.6,47.6)
Long<-c(-92.5,-92.1,-93.3,-93.2,-91.9,-95.9,-90.8)
Place.type<-c("City","City","City","Town","Town","Town","Town")

MN_Cities<-data.frame(City,Lat,Long,Place.type)

ggplot(data = mn_data, mapping = aes(x = long, y = lat)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")+ 
  geom_point(
    data = MN_Cities,
    aes(x = Long, y = Lat, color=Place.type),
    alpha = 1, size = 2)+
  geom_text(data=MN_Cities, aes(Long, Lat, label=City))

```



# Map of all Montana Starbucks 

\

Q: What do you notice about the laeling of the cities, is it helpful or not?  Do the labels tell you anything?

\
  
  
```{r}

MT_Starbucks<-starbucks%>%
  filter(State.Province=="MT")%>%
  select(Store.Number,Store.Name,Ownership.Type,Street.Address,City,Longitude,Latitude)

mt_data<-subset(states, region == "montana")

ggplot(data = mt_data, mapping = aes(x = long, y = lat)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")+ 
  geom_point(
    data = MT_Starbucks,
    aes(x = Longitude, y = Latitude, color=Ownership.Type),
    alpha = 1, size = 2)+
  geom_text(data=MT_Starbucks, aes(Longitude, Latitude, label=City))



ggplot(data = mt_data, mapping = aes(x = long, y = lat)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")+ 
  geom_point(
    data = MT_Starbucks,
    aes(x = Longitude, y = Latitude, color=Ownership.Type),
    alpha = 1, size = 2)

  
```

\

# Data of birth places for the class

\


```{r}

# Section 2 Data
# City<-c("Berkeley","Madison", "Chicago", "St. Louis Park","Yerevan", "Burnsville",
#         "Topeka","WashingtonDC","Nanjing","Chicago","Minneapolis","Rochester",
#         "Madison","Ibadan","Rockport")
# Lat<-c(37.9, 43.1, 41.9, 44.9, 40.2, 93.3, 39.0, 38.9, 32.1, 41.9, 44.5, 
#  44.0, 43.1 , 3.9, 44.2)
# Long<-c(-122.3,-89.4,-87.6,-93.4, 44.5, -44.8, -95.7,-77.0, 118.8, 
# -87.6, -93.5,-92.5, -89.4,7.4,-69.1)


# Section 3 Data
City<-c("Antioch", "Surakarta", "Santiago", "Minneapolis", "Suzhou", "Wheat Ridge", "Cleveland", "NYC", "North Platte")
Lat<-c(38, -7.6, -33.4, 45, 31.3, 39.8, 41.5, 40.7, 42)
Long<-c(121.8, 111.3, -70.7, -93.3, 120.6, -105.1, -81.7, -73.9,-100.5)

Class_Places_Data<-data.frame(City,Lat,Long)

# Class_Places_Data[6,2]<- 44.8
# Class_Places_Data[6,3]<- -93.3


ggplot(world_boundaries) +
   geom_sf() +
   geom_point(
     data = Class_Places_Data,
     aes(x = Long, y=Lat),
     alpha = 1, size = 2)+
  geom_text(data=Class_Places_Data, aes(Long, Lat, label=City))


```


# Activity

\
```{r}
states <- map_data("state")

co_data<-subset(states, region == "colroado")

# colorado Data
City<-c("Denver", "Colorado Springs", "Lakewood", "Aurora", "Boulder")
Lat<-c(39.739235, 38.833881, 39.710999, 39.712929, 40.232281)
Long<-c(-104.990250, -104.821365, -105.088867, -104.819359, -105.181389)

Class_Places_Data<-data.frame(City,Lat,Long)


ggplot(data = co_data, mapping = aes(x = long, y = lat)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")+ 
  geom_point(
    data = Class_Places_Data,
    aes(x = Long, y = Lat),
    alpha = 1, size = 2)


```
Choose the state you were born in (if you were born in Minnesota or if you were born outside the USA choose another state where you know someone). 

Do a Google search and find the 5 largest cities in that state, find the latitudes and longitudes of those cities.

Make a map of that state with the cities labeled

