---
title: "Project"
author: "Brady,Sofie,Mary"
date: "2024-04-11"
output: 
  html_document:
  toc: true
  toc_float: true
  theme: cosmo
  highlight: espresso
---

# Project Topic
  Student alcohol consumption
# One Question of interest
  How does a student's homelife impact their alcohol consumption
  
```{r}
library(tidyverse)
```

```{r}
data1 = read.csv("https://raw.githubusercontent.com/Brady3035/Comp-112/main/student-lpor.csv")
vars_of_intrest <- data1 %>%
  select(famsize,Pstatus,Medu,Fedu,Fjob,Mjob,famsup,Dalc,Walc,health,absences,age)
```

```{r}
ggplot(data = vars_of_intrest, aes(x=age,y = Dalc))+
  geom_point()+
  facet_wrap(~Pstatus) + geom_jitter()
```


Here we have a graph that adds up the categorical variables Dalc and Walc which represent on a scale of 1-5 how much alcohol a highschool student in portugal drinks on working days vs weekend days, against their average absenses to try and get an idea if more alcohol results in more absenses from school. 

``` {r,echo=FALSE}

vars_of_intrest <- mutate(vars_of_intrest, Dalc_Walc = Dalc + Walc)

mean_absences <- vars_of_intrest %>%
  group_by(Dalc_Walc) %>%
  summarise(mean_absences = mean(absences))

ggplot(mean_absences, aes(x = factor(Dalc_Walc), y = mean_absences)) +
  geom_col() +
  labs(x = "Dalc + Walc", y = "Average Absences") +
  theme_minimal()

```