---
title: "Global Starbucks Locations"
author: "Dan O'Loughlin"
date: "2024-02-06"
output: html_document
---


```{r include = FALSE}
# This chunk just sets up some styling (eg: default size of our images)
knitr::opts_chunk$set(
  collapse = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.height = 8, 
  fig.width = 8,
  fig.align = 'center')
```



\
\


### Our goal in this short Document is to make a world map and color it with the location of all the Starbucks in the world (at least at the time the data was collected)


\
\

```{r}
# Load the necessary packages
library(tidyverse)
library(rnaturalearth)

# Get info about country boundaries across the world
# in a "sf" or simple feature format
world_boundaries <- ne_countries(returnclass = "sf")


# Import starbucks location data
starbucks <- read.csv("https://bcheggeseth.github.io/112_fall_2023/data/starbucks.csv")
```

\


In your **console**, type `world_boundaries` to see what is stored there.
Don't print it our in your Rmd -- printing it would be really messy there (even just the `head()`).

then, also in the **console** run the command

`starbucks <- read.csv("https://bcheggeseth.github.io/112_fall_2023/data/starbucks.csv")`

and in the **console** and type **View(starbucks)**

to see what is stored in that dataset.

\
\


- **Make a blank map**

\


```{r}
# What does this code produce?
# What geom are we using for the point map?
ggplot(world_boundaries) + 
  geom_sf()
```

\


- **add the locations of the Starbucks stores using the Lat and Long stored in the `starbucks` dataset for each store**

\



```{r}
# Load package needed to change map theme
library(mosaic)

# Now add a point for each Starbucks
# NOTE: The Starbucks info is in our starbucks data, not world_boundaries
# How does this change how we use geom_point?!
ggplot(world_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3, size = 0.2, color = "darkgreen"
  ) +
  theme_map()
```



