---
title: "Visualization Practice"
author: "Dan O'Loughlin"
date: "2024-02-08"
output: 
  html_document:
    toc: true
    toc_float: true
---



```{r include = FALSE}
# This chunk just sets up some styling (eg: default size of our images)
knitr::opts_chunk$set(
  collapse = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.height = 4, 
  fig.width = 6,
  fig.align = 'center')
```


# Practice with visualization

Using a dataset called `SaratogaHouses` we will do some practice with visualization. 

\



```{r}

# load packages including data

library(tidyverse)
library(mosaicData)

## SaratogaHouses is Data on houses in Saratoga County, New York, USA in 2006
## it is part of the mosaicData package

head(SaratogaHouses)

str(SaratogaHouses)

```

\

# Plot No. 1

\


Consider the code below which 

- Makes a plot showing three variables

- Scales one of the variables to make an axis more readable

- Makes a new plot with the newly scaled variable that also has labels



\


```{r}

ggplot(SaratogaHouses,aes(x=livingArea,y=price,color=centralAir))+
  geom_point()+
  geom_smooth(se=FALSE,method=lm)

## NOTE:  the following command is not optimal since it changes the data
## we wil learn better ways to do change the scale of a variable in the next unit
## on Data wrangling

SaratogaHouses$price<-SaratogaHouses$price/1000

ggplot(SaratogaHouses,aes(x=livingArea,y=price,color=centralAir))+
  geom_point()+
  geom_smooth(se=FALSE,method=lm)+
  labs(color="Central Air",title="square footage of homes vs price by central air in Saratoga")+
  xlab("square footage")+ylab("price in 1K dollars")

```

## Questions about plot 1

\

- What does this plot tell us about these three variables?

- What are some other plots we could make which could help illuminate the relationship between these three variables?

\

# Activity
```{r}
ggplot(SaratogaHouses,aes(x=livingArea,y=price,color = newConstruction))+
  geom_point()+
  geom_smooth(se=FALSE,method=lm)+
  labs(color="New Construction",title="square footage of homes vs price by type of heating colored by new constrction status in Saratoga")+
  xlab("square footage")+ylab("price in 1K dollars") + 
  facet_wrap(~fuel)

```
\

## No 1.  

Write code below which makes one more plot using these three variables, then say what your plot shows. Your plot should include labels (you might want to look at help("labs"))

\

**Put your code here**

\
\


## No 2.  

Write code below which makes a plot using three different variables from the dataset. Again, say what your plot shows, and your plot should include labels.

\

**Put your code here**
```{r}
ggplot(SaratogaHouses,aes(x=bathrooms))+
  geom_density()+
  labs(title="Bathroom number distibution given bedroom count")+
  xlab("Bathrooms")+ylab("Count") +
  facet_wrap(~bedrooms)
```
